---
title: "US Presidential Popular Vote Margins"
excerpt: "Since 1976 who's won big?"
date: '2020-11-08'
draft: no
showpagemeta: yes
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
```

Let's look at the margin of victory of US presidents since 1976. Here we find the 
vote count for the democratic and republican candidate each year, and calculate the 
percentage difference between the two (such that the democratic victory 
`dem_perc_winloss` is postive.)

We pull the [U.S. President 1976–2016 dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/42MVDX)
from the [MIT election lab](https://electionlab.mit.edu/data).

```{r message=FALSE, warning=FALSE, include=FALSE}
df <- read_csv('../data/1976-2016-president.csv')
```


```{r include=FALSE}
df_before_2020 <- df %>%
  # In Minnesota, Ds run under weird party name
  mutate(party = ifelse(party == 'democratic-farmer-labor', 'democrat', party),
         party = ifelse(party == 'democratic', 'democrat', party)) %>%  
  filter(party %in% c('republican','democrat'),
         !str_detect(candidate,'Gritz, James'), # Ran as democrat/republican
         writein == FALSE) %>%
  group_by(year, candidate, party) %>%
  summarize(candidate_votes = sum(candidatevotes)) %>% 
  pivot_wider(names_from=party, values_from=c(candidate, candidate_votes))
```

For the 2020 election, let's add in results from the
[WSJ](https://www.wsj.com/election-results-2020/) as of Nov 8 2020. At that time
it showed 75.2M votes for Joe Biden and 70.8M votes cast for Donald Trump.

```{r echo=FALSE}
library(knitr)
df_2020 <- tibble(year = 2020,
                  candidate_democrat = 'Joe Biden',
                  candidate_republican = 'Donald Trump',
                  candidate_votes_democrat = 75215986,
                  candidate_votes_republican = 70812803) 

d_combined <- dplyr::bind_rows(df_before_2020, df_2020)

d_combined %>%
  mutate(dem_perc_winloss = 100 * ((candidate_votes_democrat - 
                                    candidate_votes_republican) / 
                                       candidate_votes_democrat)) %>% 
  # Make cols names more readable
  rename(dem_candidate = candidate_democrat,
         repub_candidate = candidate_republican,
         dem_votes = candidate_votes_democrat,
         repub_votes = candidate_votes_republican) %>% 
  knitr::kable()
```

Overall, it's amazing how the vote margins have shrunk compared to elections in the
80s and 90s. **In terms of the current election, it appears that Joe Biden has achieved
a larger popular vote victory than any Republican president since George HW Bush's 
victory in 1988, 32 years ago.**
