---
title: "US presidential popular vote margins"
date: '2020-11-08'
draft: no
showpagemeta: yes
excerpt: "Since 1976 who's won big?"
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dev = 'png',
                      dpi = 200, out.width='650px', fig.width=8, fig.height=4)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
```

Let's look at the margin of victory of US presidents since 1976. Here we find the 
vote count for the democratic and republican candidate each year, and calculate the 
percentage difference between the two.

We first pull the [U.S. President 1976–2016 dataset](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/42MVDX)
from the [MIT election lab](https://electionlab.mit.edu/data).

```{r message=FALSE, warning=FALSE, include=FALSE}
df <- read_csv('../data/1976-2016-president.csv')
```


```{r include=FALSE}
df_before_2020 <- df %>%
  # In Minnesota, Ds run under weird party name
  mutate(party = ifelse(party == 'democratic-farmer-labor', 'democrat', party),
         party = ifelse(party == 'democratic', 'democrat', party),
         candidate = ifelse(candidate == 'Obama, Barack H.', 'Obama, Barack', candidate)) %>% 
  filter(party %in% c('republican','democrat'),
         !str_detect(candidate,'Gritz, James'), # Ran as democrat/republican
         writein == FALSE) %>%
  group_by(year, candidate, party) %>%
  summarize(candidate_votes = sum(candidatevotes)) %>% 
  pivot_wider(names_from=party, values_from=c(candidate, candidate_votes))
```

For the 2020 election, let's add in results from the
[WSJ](https://www.wsj.com/election-results-2020/) as of Nov 8 2020. At that time
it showed 75.2M votes for Joe Biden and 70.8M votes cast for Donald Trump.

The resulting table shows the raw vote totals as well as the percentage difference
(such that the democratic victory are positive in the `dem_perc_win` 
column below). *Note, this isn't percentage of total votes, but percentage between dems and repub candidates. Also note that raw values mean little because of population growth.*

```{r include=FALSE}
library(knitr)
df_2020 <- tibble(year = 2020,
                  candidate_democrat = 'Joe Biden',
                  candidate_republican = 'Donald Trump',
                  candidate_votes_democrat = 75215986,
                  candidate_votes_republican = 70812803) 

d_combined <- dplyr::bind_rows(df_before_2020, df_2020)
```

```{r echo=FALSE}
library(kableExtra)
d_combined %>%
  mutate(dem_perc_win = round(100 * ((candidate_votes_democrat - 
                                    candidate_votes_republican) / 
                                       candidate_votes_democrat), 2)) %>% 
  # Make col names more readable
  rename(dem_candidate = candidate_democrat,
         repub_candidate = candidate_republican,
         dem_votes = candidate_votes_democrat,
         repub_votes = candidate_votes_republican) %>% 
         kable() %>% 
         kable_styling(bootstrap_options = c("striped", "hover", 
                                             "condensed", "responsive"))
```

Overall, it's amazing Ronald Reagan won the popular vote by ~45% in 1984. **In terms of the current election, it appears that last Republican president to achieve a larger popular vote victory than Joe Biden was George HW Bush in 1988, 32 years ago.**
